---
title: "Demographic Changes"
author: "Karolina DM"
output: html_document
---

```{r setup, include=FALSE}
library("knitr")
library("tidyverse")
library("plotly")
library("htmlwidgets")
source("DemographicChanges.R")
```


## This is a report about demographic changes in the world since 1960 to 2019. For this report I choose selected demographic factors, like:

- __Fertility Rate__,
- __Birth Rate__,
- __Death Rate__,
- __Life Expectancy__


***

__Fertility Rate__ (total, births per woman) is:

> "rate represents the number of children that would be born to a woman if she were to live to the end of her childbearing years and bear children in accordance with age-specific fertility rates of the specified year".

(source of definition & data: https://data.worldbank.org/indicator/SP.DYN.TFRT.IN)


The country with the highest fertility rate in 2019 is 
__`r HighFertRat$Country.Name`__. The fertility rate is 
__`r HighFertRat$hfr`__.

The country with the lowest fertility rate in 2019 is 
__`r LowFertRat$Country.Name`__. The fertility rate is 
__`r LowFertRat$lfr`__.

This is the list of countries with the highest fertility rate (ordered by highest to lowest). Table shows fertility rate difference declining from 1960 to 2019.

```{r top_10_hfr, echo=FALSE}
kable(top_10_hfr, col.names = c("Country", "Fertility Rate drop since 1960 to 2019"))
```

This map shows fertility rate in 2019. 

```{r , echo=FALSE}
mapCountryData(
  comb_data_map_fr,
  mapTitle = "Fertility Rate in 2019",
  nameColumnToPlot = "X2019",
  colourPalette = brewer.pal(7, "RdPu")
)
```

Fertility rate drops globally. 

The country of EU with the highest fertility rate in 2019 is 
__`r EU_HighFertRat$Country.Name`__. The fertility rate is 
__`r EU_HighFertRat$eu_hfr`__.
The high fertility rate in France is the result of counting a combined result with dependent territories in the whole world (e.g. French Guiana, French Polynesia, Martinique).

The country of EU with the lowest fertility rate in 2019 is 
__`r EU_LowFertRat$Country.Name`__. The fertility rate is 
__`r EU_LowFertRat$eu_lfr`__.

This map shows fertility rate in EU countries in 2019. 

```{r , echo=FALSE, out.width="150%" }
mapCountryData(
  comb_data_map_eu_fr,
  mapTitle = "Fertility Rate in 2019 in EU",
  nameColumnToPlot = "X2019",
  colourPalette = brewer.pal(7, "RdPu")
)
```

This chart shows fertility rate in EU countries in 2019.

```{r EU_FR=plot, dev='png', out.width="150%", echo=FALSE}
plot(EU_FR)
```

There are many correlates of the fertility rate. One of them is the age of the mother when she gives birth of her first child. 
Correlation between fertility rate and mean age of woman at birth of first child is
__```r cor(comb_data_agewom_fr$X2019, comb_data_agewom_fr$OBS_VALUE, use = "complete.obs")```__
The higher the age of birth of first child, the lower the fertility rate.

This chart shows fertility rate and mean age of woman at birth of first child (in EU countries in 2019).

```{r ggFR=plot, dev='png', out.width="150%", echo = FALSE}
plot(ggFR)
```

As we can see in Eastern European countries belonging to the EU, the fertility rate is higher and the age of women giving birth to the first child is lower than in other EU countries.

One of the potential correlates of fertility rate and the age at which women give birth to first child may be a greater prevalence of contraception and therefore - more frequent use of it.
Correlation between fertility rate and prevalence of contraceptive methods is
__```r cor(comb_data_agewom_fr_contr$X2019, comb_data_agewom_fr_contr$AnyMethod, use = "complete.obs")```__ This correlation is very weak. 

(Source of data about prevalence of contraceptive methods in EU countries: https://www.un.org/development/desa/pd/sites/www.un.org.development.desa.pd/files/files/documents/2020/Jan/un_2019_contraceptiveusebymethod_databooklet.pdf)


This interactive graph shows the distribution of the variables: the fertility rate, the average age of the first child's birth, and the prevalence of contraception used (hover over a dot on chart for more information).

```{r plt=ggplotly, out.width="110%", echo=FALSE, error=FALSE, warning=FALSE}
ggplotly(plt)
```


Another correlates of fertility rate could be GCP per capita, which is:

> "The indicator is calculated as the ratio of real GDP to the average population of a specific year. GDP measures the value of total final output of goods and services produced by an economy within a certain period of time. It includes goods and services that have markets (or which could have markets) and products which are produced by general government and non-profit institutions. It is a measure of economic activity and is also used as a proxy for the development in a country’s material living standards. However, it is a limited measure of economic welfare. For example, neither does GDP include most unpaid household work nor does GDP take account of negative effects of economic activity, like environmental degradation".

(data & definition source: https://ec.europa.eu/eurostat/databrowser/view/sdg_08_10/default/table).

We want to check if GDP correlates with the factors, which we discussed earlier.

Correlation between __fertility rate__ and __GDP__ is
__```r cor(comb_data_agewom_fr_contr_gdp$X2019,comb_data_agewom_fr_contr_gdp$GDP_per_cap)```__ This correlation is really weak. But SD of fertility rate is low:  
__```r sd(comb_data_agewom_fr_contr_gdp$X2019)```__ , what it means, that the values observed are centered around the mean value.

Correlation between __mean age at the birth of first child__ and __GDP__ is
__```r cor(comb_data_agewom_fr_contr_gdp$OBS_VALUE,comb_data_agewom_fr_contr_gdp$GDP_per_cap)```__ This is strong correlation. 

The situation is similar in the case of the next correlated pair of variables: __contraceptive methods__ and  __GDP__. Correlation is
__```r cor(comb_data_agewom_fr_contr_gdp$AnyMethod, comb_data_agewom_fr_contr_gdp$GDP_per_cap, use = "complete.obs")```__ and  also is strong. 

This graph shows the correlations between the variables discussed above. 

``` {r , out.width="150%", echo=FALSE, error=FALSE, warning=FALSE}
ggpFR
```

As we can see...


Birth Rate (crude, per 1000 people) is: 

>Crude birth rate indicates the number of live births occurring during the year, per 1,000 population estimated at midyear. Subtracting the crude death rate from the crude birth rate provides the rate of natural increase, which is equal to the rate of population change in the absence of migration.

(data & definition source: https://data.worldbank.org/indicator/SP.DYN.CBRT.IN)

The country with the highest birth rate in 2019 is 
`r HighBirthRat$Country.Name`. The birth rate is 
`r HighBirthRat$hbr`.

The country with the lowest birth rate in 2019 is 
`r LowBirthRat$Country.Name`. The birth rate is 
`r LowBirthRat$lbr`.

This is the list of countries with the highest birth rate (ordered by highest to lowest). Table shows birth rate difference declining from 1960 to 2019.

```{r top_10_hbr, echo=FALSE}
kable(top_10_hbr, col.names = c("Country", "Birth Rate drop since 1960 to 2019"))
```

This map shows birth rate in 2019. 

```{r , echo=FALSE}
mapCountryData(
  comb_data_map_br,
  mapTitle = "Birth Rate in 2019",
  nameColumnToPlot = "X2019",
  colourPalette = brewer.pal(7, "Greens")
)
```

This chart shows birth rate in the whole world in 2019 divided into continents.
``` {r , out.width="150%", echo=FALSE, error=FALSE, warning=FALSE}
BR_cont
```

The highest birth rate in 2019 is observed on the African continent, and the lowest - in Europe. This is also confirmed by the calculations.

```{r calcBR, echo=FALSE}
kable(calcBR, col.names = c("Continent", "Mean", "Median", "SD"), caption = "Birth rate in 2019 - selected statistics")
```

__Death Rate__ (crude, per 1000 people) indicates: 

> "the number of deaths occurring during the year, per 1,000 population estimated at midyear. Subtracting the crude death rate from the crude birth rate provides the rate of natural increase, which is equal to the rate of population change in the absence of migration".

(data & definition source: https://data.worldbank.org/indicator/SP.DYN.CDRT.IN)

The country with the highest death rate in 2019 is 
__`r HighDeathRat$Country.Name`__. The death rate is 
__`r HighDeathRat$hdr`__.

The country with the lowest death rate in 2019 is 
__`r LowDeathRat$Country.Name`__. The death rate is 
__`r LowDeathRat$ldr`__.

This is the list of countries with the highest death rate (ordered by highest to lowest). Table shows death rate difference increasing from 1960 to 2019.

```{r top_10_hdr, echo=FALSE}
kable(top_10_hdr, col.names = c("Country", "Death Rate increase since 1960 to 2019"))
```

This map show death rate in 2019. 

```{r , echo=FALSE}
mapCountryData(
  comb_data_map_dr,
  mapTitle = "Death Rate in 2019",
  nameColumnToPlot = "X2019",
  colourPalette = brewer.pal(7, "Blues")
)
```

This boxplot shows death rate in 2019.

```{r boxplot_DR=plot, dev='png', out.width="150%", echo = FALSE, warning=FALSE}
plot(boxplot_DR)
```

The highest death rate in 2019 is observed on the European continent, and the lowest - in Africa This is also confirmed by the calculations.

```{r calcDR, echo=FALSE}
kable(calcDR, col.names = c("Continent", "Mean", "Median", "SD"), caption = "Death rate in 2019 - selected statistics")
```


The country with the highest death rate in 2019 in EU is 
__`r EU_HighDeathRat$Country.Name`__. The death rate is 
__`r EU_HighDeathRat$eu_hdr`__.

The country with the lowest death rate in 2019 is 
__`r EU_LowDeathRat$Country.Name`__. The death rate is 
__`r EU_LowDeathRat$eu_ldr`__.


This map show death rate in 2019 in EU. 

```{r , echo=FALSE}
mapCountryData(
  comb_data_map_eu_dr,
  mapTitle = "Death Rate in 2019 in EU",
  nameColumnToPlot = "X2019",
  colourPalette = brewer.pal(7, "Blues")
)
```

Reasons for increasing the death rate (in EU in 2017) are:

- circulatory disease (37%),
- cancer(25%),
- other (17%) like: mental and behavioral disorders, endocrine, nutriitional and metabolic deseases, and all remaining causes,
- respiratory diseases (8%),
- external causes of death (5%),
- diseases of digestive system (4%),
- diseases of the nervous system (4%)

(source: https://ec.europa.eu/eurostat/statistics-explained/index.php?title=Causes_of_death_statistics#Major_causes_of_death_in_the_EU_in_2017)

Animation shows death rate in one of the EU country - Poland since 1960 to 2019:

``` {r , out.width="150%", echo=FALSE, error=FALSE, warning=FALSE}
ggPL_DR
```

As we can see death rate in Poland increases since 1960. Moreover, the current pandemic situation is likely to have a significant impact on the above indicator.


Life expectancy at birth (total, years): 

>indicates the number of years a newborn infant would live if prevailing patterns of mortality at the time of its birth were to stay the same throughout its life.

(data & definition source: https://data.worldbank.org/indicator/SP.DYN.LE00.IN)

The country with the highest life expectancy in 2019 is 
`r HighLifeExpect$Country.Name`. The life expectancy rate is 
`r HighLifeExpect$hle`.

The country with the lowest life expectancy in 2019 is 
`r LowLifeExpect$Country.Name`. The life expectatancy is 
`r LowLifeExpect$lle`.

This is the list of countries with the largest life expectancy gains from 1960 to 2019

```{r top_10_hleg, echo=FALSE}
kable(top_10_hleg, col.names = c("Country", "The greatest increase since 1960 to 2019 (in years)"))
```

This map show life expectancy in 2019 in the whole world. 

```{r , echo=FALSE}
mapCountryData(
  comb_data_map_le,
  mapTitle = "Life expectancy in 2019",
  nameColumnToPlot = "X2019",
  colourPalette = brewer.pal(7, "Purples")
)
```

The factors which have impact for life expectancy are among others:

- income per capita, the rate of hospital beds, medical staff (Cervantes, López & Rambaud, 2019);

- universal health coverage, child vaccination, sanitary conditions (Ranabhat et all., 2018);

- real income per capita, educational level (Keneci & Murthy, 2017);

- economic growth, health expenditure, clean water, improved sanitation conditions, carbon emission, access to drinking water (Rahman, Rana & Khanam, 2022)


The country of EU with the highest life expectancy in 2019 is 
__`r EU_HighLE$Country.Name`__. The life expectancy is 
__`r EU_HighLE$eu_hlifexp`__.

The country of EU with the lowest life expectancy in 2019 is 
__`r EU_LowLE$Country.Name`__. The fertility rate is 
__`r EU_LowLE$eu_llifexp`__.

This map shows life expectancy in EU countries in 2019. 

```{r , echo=FALSE, out.width="150%" }
mapCountryData(
  comb_data_map_eu_le,
  mapTitle = "Life Expectancy in 2019 in EU",
  nameColumnToPlot = "X2019",
  colourPalette = brewer.pal(7, "Purples")
)
```

As we can see, the citizens of the Mediterranean basin, countries of the western and northern parts of the European Union achieve the longest life expectancy.

Mental factors can also affect life expectancy. Therefore, it was checked whether depressive symptoms correlate with life expectancy in the EU countries. Correlation between __life expectancy__ and __depressive symptoms__ is
__```r cor(EU_LifeExpect_depr_GDP$X2019, EU_LifeExpect_depr_GDP$Depr_sympt, use = "complete.obs") ```__ It is a weak correlation.

This graph shows the correlations between the life expectancy, GDP (per capita) and depressive symptoms. 

``` {r , out.width="150%", echo=FALSE, error=FALSE, warning=FALSE}
ggpLE
```

Correlation between __life expectancy__ and __GDP__ is
__```r cor(EU_LifeExpect_depr_GDP$X2019, EU_LifeExpect_depr_GDP$GDP_per_cap, use = "complete.obs")```__ - it is a strong correlation.